Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-08-30 alloscore_manuscript/manuscript.Rnw
typo fix in joint dependence comment

We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (cf. equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix). It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of their forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.  In this sense, the AS depends “jointly” on the marginal forecasts while the WIS does not.

@@

q1 <- expexamp$Q[[1]]
q2 <- expexamp$Q[[2]]
q_sum <- function(tau) q1(tau) + q2(tau)
iters5 <- tibble(taus = 1-alloc$lam_seq[[1]], qtot = q_sum(taus))
iters10 <- tibble(taus = 1-alloc$lam_seq[[2]], qtot = q_sum(taus))

@@

p5 <- p + geom_hline(aes(yintercept = 5)) +
  geom_hline(aes(yintercept = c(1,4)), linetype = 2) +
  geom_segment(data = iters5, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,0.632,.7,.8,.9,1),
                     labels = c("0.5","0.6",expression(italic(atop(" ", 1 - e^{-1} %~~% 0.632))),
                                "0.7","0.8","0.9","1")) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,1,4,5,10,15), expand = c(0,0))

p10 <- p + geom_hline(aes(yintercept = 10)) +
  geom_hline(aes(yintercept = c(2,8)), linetype = 2) +
  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot), color = "red", alpha = .5) +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,.7,.8,0.865,.9,1),
                     labels = c("0.5","0.6",
                                "0.7","0.8", 
                                expression(italic(atop(" ", 1 - e^{-2} %~~% 0.865)), "0.9", "1"))) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 
                     name = "  ")

cowplot::plot_grid(p5, p10, align = "h")

>>>>

We also note that the AS, just as the (aggregate or mean) quantile score and WIS, depends only on the marginal distributions for each location of a multi-location forecast (see equations \eqref{eqn:AS-raw} and \eqref{eqn:quantiles-sum-to-K} in Section \ref{sec:a:bayes-quantiles} of the appendix). It is however the case that when a forecaster modifies a single marginal forecast (holding those for other locations fixed), the aggregate WIS of their forecast will change only according to the WIS of that marginal forecast, whereas every term in the sum defining the AS (in \eqref{eqn:AS-raw}) will change as the Bayes probability level defined by \eqref{eqn:quantiles-sum-to-K} changes.  In this sense, the AS depends “jointly” on the marginal forecasts while the WIS does not.

@@

q1 <- expexamp$Q[[1]]
q2 <- expexamp$Q[[2]]
q_sum <- function(tau) q1(tau) + q2(tau)
iters5 <- tibble(taus = 1-alloc$lam_seq[[1]], qtot = q_sum(taus), ind = seq_along(alloc$lam_seq[[1]]))
iters10 <- tibble(taus = 1-alloc$lam_seq[[2]], qtot = q_sum(taus), ind = seq_along(alloc$lam_seq[[2]]))

@@

p5 <- p + geom_hline(aes(yintercept = 5)) +
  geom_hline(aes(yintercept = c(1,4)), linetype = 2) +
  geom_segment(data = iters5, aes(x=taus, xend = taus, y=0, yend = qtot, color = ind), alpha = .5) +
  scale_color_viridis_c("Iteration\n index", end = 0.75, direction=-1, limits = c(1, 18), option="B") +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,0.632,.7,.8,.9,1),
                     labels = c("0.5","0.6",expression(italic(atop(" ", 1 - e^{-1} %~~% 0.632))),
                                "0.7","0.8","0.9","1")) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,1,4,5,10,15), expand = c(0,0)) +
  theme(legend.position = "none")

p10 <- p + geom_hline(aes(yintercept = 10)) +
  geom_hline(aes(yintercept = c(2,8)), linetype = 2) +
  geom_segment(data = iters10, aes(x=taus, xend = taus, y=0, yend = qtot, color = ind), alpha = .5) +
  scale_color_viridis_c("Iteration\nindex", end = 0.75, direction=-1, limits = c(1, 18), option="B") +
  scale_x_continuous(limits = c(.45,1), expand = c(0,0), 
                     breaks = c(.5,.6,.7,.8,0.865,.9,1),
                     labels = c("0.5","0.6",
                                "0.7","0.8", 
                                expression(italic(atop(" ", 1 - e^{-2} %~~% 0.865)), "0.9", "1"))) +
  scale_y_continuous(limits = c(0,12), breaks = c(0,2,5,8,10,15), expand = c(0,0), 
                     name = "  ")

legend <- get_legend(p10 + theme(legend.box.margin = margin(0, 0, 0, 12)))
p10 <- p10 + theme(legend.position = "none")

plot_grid(
  cowplot::plot_grid(p5, p10, align = "h"),
  legend,
  rel_widths = c(3, 0.4)
)
@
